---
title: "CSC8631 Report"
author: "Thomas Richardson"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir= normalizePath('..'))
```

```{r ProjectTemplate, include=FALSE}
library("ProjectTemplate")
load.project()
source("src/eda.R")
```

# Cycle 1
## Business Understanding
FutureLearn is a Massive Online Open Course (MOOC) platform, and since launching
in 2013 they have partnered with a range of universities and businesses to
provide a variety of online courses and degrees (FutureLearn: Online Courses and
Degrees from Top Universities). They are situated in the Higher Education space
and are hence competing with a number of new entrants in the marketplace many of
which have an emphasis on attracting consumers in the international market 
(Shacklock, X., 2016). In order to gain an advantage in this area, FutureLearn 
are looking to incorporate Learning Analytics to both improve their courses and 
develop a strategy to attract more consumers. The course being investigated for
this project is a Cyber Security course led by Newcastle University with 
specific success criteria of finding an accessible method to encourage consumers
to sign up to a FutureLearn course and determining methods to adjust the course 
infrastructure so that these people feel more satisfied throughout.

The first data mining goal for this investigation was to identify at which times
throughout the year the course was most popular in an attempt to gain an
understanding of why this may be the case and allow for future planning so
that increased support can be made available to the participants to help keep up
with the increasing demand during these periods. In addition to this a 
by-product of this investigation may enable FutureLearn to identify periods to 
focus their advertising budget, in an attempt to optimise their resources and
potentially gain an advantage over their competition. If successful there will
be clear recurring time period(s) of increased engagement that can be seen from
the data.

A constraint to the investigation is that there was no data in 
regard to previous advertising of the course and how this may have affected 
enrolments so for this reason it was assumed throughout that advertising had
remained at a constant level spread evenly across all regions and mediums for 
the entirety of the course's lifespan. In addition, it was to be assumed that the
course was available to enrol upon at moment throughout the course history.

This project follows the CRISP-DM process model (Chapman et al., 2000) and all 
data processing was completed in R, incorporating the package `ProjectTemplate`.
Other R libraries used include `ggplot2` which was to be used to create plots 
due to its flexibility in creating plots allowing for ease of use when layering
and using position adjustment in plots, and `dplyr` which was to be used to
transform the data in an efficient manner.

## Data Understanding
As previously mentioned, the data collected was from the Newcastle University
Cyber Security MOOC. It was collected over 7 consecutive runs of the course, with
each run spanning a different length of time ranging from roughly two months in 
duration to over sixth months. There were often slight changes made to the
course between runs and there were also sometimes some data collection changes
between runs, for example data collection on team members did not start until
the second run and hence this csv file for the first run does not exist.

The data had been grouped into (up to) eight different csv files per run,

- **Archetype Survey Responses:** This contained learner IDs, the time and
date they completed the survey, and the archetype they were given based on their
results.
- **Enrolments:** This contained learner IDs, the date and time they enrolled,
unenrolled, completed the course and purchased a certificate. It also contains
information such as the individuals' roles on the course (e.g., learner), their age
range, gender, country, highest education level, employment status and 
employment area, and finally the country they were detected from.
- **Leaving Survey Responses:** This contained learner IDs, the date and time
they left the course as well as details on their last completed step, and a
reason for leaving.
- **Question Responses:** This contained learner IDs, information to identify 
the question (e.g., question number), the type of question (e.g., multiple 
choice), the learners' answers and whether they answered correctly.
- **Step Activity:** This contained learner IDs, and when they first visited and
last completed each step.
- **Team Members:** This contained information on the different team members such
as their team role and user role.
- **Video Stats:** This contained statistics on each of the videos on the course 
including video length, views, percentage of viewers who reached different
percentage lengths of the videos, the proportion of device types used to watch
and the proportion of viewers per continent.
- **Weekly Sentiment Survey Responses:** This contained feedback responses 
which consisted of the week they were submitted, an experience rating and a
reason for that rating.

Due to the differing lengths of each of the runs there were some runs with
significantly less data than others and due to learners dropping out at various
points throughout the course there was henceforth less data on these later 
stages. Not all of the data collected was quantitative, as data such as feedback
also takes qualitative responses and while these can offer greater insights into
learner attitudes and opinions they also allow for brief non-descriptive 
opinions such as a response found in the **Weekly Sentiment Survey Responses**
of the seventh run which was simply "DEAD TING". In spite of this the overall
scope of the data was wide and allowed for a variety of investigative approaches
in a variety of areas.

As per the goals of this investigation it appeared that the most useful data 
would be found in the **Enrolment** files across all 7 of the runs as from this 
it would be possible to investigate the number of learners enrolled since the
course began and at which points in time the number of enrolments was at a 
greater level than others.

## Data Preparation
Following on from the data understanding the first act to performed on the data
was to merge the 7 **Enrolment** files from the different runs. This was done
using the `merge` function in R and can be seen in the 01-A.R file in the munge
folder of the FutureLearn repository. It was felt important to use data from all
7 of the runs so as to investigate across as much data as possible in order to
gain more reliable results. Following this the data was cut down to only contain
the rows where the learner's role was labelled as "learner" so as to not
potentially skew results by including data of other roles. Next, making use of
the `dplyr` package, a new data frame was made, transforming the data so that this
new data frame contained a count of the number of enrolments in each month of the 
course. Months that contained no enrolments were not included in this data frame but
were to be accounted for on future graphs. Finally, another column was added to
the data frame that, with help from the `lubridate` package, contained the date of
the first day of each of the months so as to identify a position on an axis for
this data to be plotted.

A similar investigation could have been performed but instead using data only 
from the learners who completed to course in order to gain an understanding of
who may most suit this course at particular times however after a short 
inspection it was clear the number of completions was not at a level high enough
to infer any reliable conclusion and for this reason enrolments remained the
focus.

The next task was to plot the data so that any potential trends or patterns 
could be seen. This was done using the `ggplot2` package to plot the bar chart
seen in Figure 1. The code for this can be found in the eda.R file in the src
folder of the FutureLearn repository.
``` {r months, fig.align="center", fig.cap="Bar chart showing the number of enrolments in each month from Ferbruary 2016 to January 2019", echo=FALSE}
months_plot
```

## Evaluation
The bars in Figure 1 appeared to follow an undulating pattern
with the highest number of enrolments to be in September 2016. There were 
common lows in the months of April and May in 2017 and 2018, and there were
also a low number of enrolments in October and November 2016.

In regard to the data mining goals, the knowledge that there exist periods of 
time both when enrolments were high and when they were low, helped to reach 
these goals however without further investigation it was difficult to make any 
concrete assumptions. It is for this reason that progression towards the 
business success criteria was also not at its most potential however this
result did leave opportunity for a directed second cycle of investigation.

Focusing on the undulating pattern, this suggested that there may be a potential
seasonal pattern to the enrolment numbers and therefore this was to be 
investigated in the succeeding cycle. In addition to this the fact that there
were dipping in enrolment numbers in the same months of consecutive year
suggests this pattern may align with global seasonal change for example, the
UK's seasons.

# Cycle 2
## Business Understanding
Due to the business objectives and success criteria not being achieved in the
first cycle a second cycle of analysis was conducted. The business objectives
and success criteria largely stayed consistent with the first cycle however with
the knowledge of a potential pattern, FutureLearn would also like to gain some
idea on why this pattern may have been occurring.

The data mining goals for this cycle carried some of the key themes from the 
previous cycle, although more concise as the goals were now to identify if the
rise and fall of enrolment numbers aligned with that of the changing of the
UK's seasons. In order to investigate this an additional assumption had to be
made. This assumption was that the UK's seasons changed around the time of
the equinoxes and solstices. That is to say that every year, at the spring
equinox (20<sup>th</sup> March) the season changes from winter to spring, at the 
summer solstice (21<sup>st</sup> June) the season changes from spring to summer,
at the autumn equinox (22<sup>nd</sup> September) the season changes from summer
to autumn, and at the winter solstice (21<sup>st</sup> December) the season
changes from autumn to winter. The success of this data mining would be judged
by the ability to form sound conclusions about whether the pattern of enrolment
numbers aligns or does not align with that of the UK's seasons.

## Data Understanding
The data did not already have the seasons allocated to the enrolment, however
as the enrolments were dated this was possible to assign during the data 
preparation. In addition, the data did not span across an exact whole number of
years and so in order to maintain a consistent number of each season across the
investigation a two-year window was selected (1<sup>st</sup> May 2016 - 
1<sup>st</sup> May 2018). This was the largest number of whole years that could
be extracted from the data and so was chosen to increase reliability.

## Data Preparation
The transforming of data for this cycle can be found in the 02-A.R file in the
munge folder of the FutureLearn repository. For this cycle a column indicating
which season the enrolment took place had to be added to the original data frame
that contained the enrolment data on all over the learners from the seven runs.
The assignment of the seasons was done by creating a function that examines a
date and returns the season based on the previously mentioned equinoxes and
solstices. This function can be found in helpers.R file in the lib folder.

Next, a new data frame was created by filtering the enrolments from the larger 
original data frame to those that fell in the selected two-year time period, and
then grouping the enrolments by season and performing a count on the numbers in
each season.

Now the data was ready to be plotted and was done so, again using the `ggplot2` 
package, to plot the bar chart seen in Figure 2. The code for this can again
be found in the eda.R file in the src folder of the FutureLearn repository.
``` {r seasons, fig.align="center", fig.cap="Bar chart showing the number of enrolments per season from 1st May 2016 to 1st May 2018", echo=FALSE}
seasons_plot
```

## Evaluation
From looking at Figure 2 it was clear to see the undulating pattern seen from
the first cycle with peaks in summer and winter (with the peak in summer being
extremely prominent), and troughs in spring and autumn. It could be suggested
from this that FutureLearn and Newcastle University could prepare for these
increased and decreased numbers of learners by adjusting their staff counts for
the course depending on the time of year so that regardless of when they are
completing the course, each learner has a sufficient level of staff support. In
addition to this there may be value to FutureLearn allocating a larger 
proportion of their advertising budget towards attracting more consumer in these
busier periods. Putting these results into context it was suggested that perhaps
these seasonal patterns were due to the breaks people are allocated in these
seasons for example summer holidays. However, due to the course being available
internationally, this raised a question of "how do these results differ
depending on the location of the individual enrolled on the course?" This
question was raised with the knowledge that different parts of the globe 
experience the seasons at different periods of time and for this a new cycle was
to be undertaken.

# Cycle 3
## Business Understanding
Although the general business aims of identifying a method to increase learner
satisfaction and to increase the total number of enrolments have been met
through the suggested plans of allocating of staff numbers depending on
enrolment levels to accommodate for demand and allocate advertising funds to 
prepare for busier periods respectively, the question raised at the end of the
previous cycle drew attention to the idea that these plans can be potentially
refined and improved upon by taking location in to consideration.

Therefore the data mining goal of this cycle was to identify if there were any
significant differences between enrolment demographics at different time periods
through the year and then in turn use these results to potentially refine the
previous recommendations to FutureLearn.

The demographics being investigated in this cycle will be geographic, more
specifically continental. This was chosen under the assumption that countries in
the same continent will be close enough in geographical distance to experience
the different seasons at the roughly the same time.

## Data Understanding
While the data did not directly specify in which continent each learner was 
located, there was a column that listed a detected country in the form of an
ISO 3166-1 alpha-2 code. Using this a continent was assigned to the individual
during the data preparation. In some instances, there was no detected country
and so these individuals will have an "NA" in their continent column. It is
also assumed for this investigation that all detected countries are correct,
however there is an understanding that some learners may be completing the
course using a VPN and hence their detected location may be inaccurate.

## Data Preparation
Prior to assigning each learner a continent, a vector was created for each
continent containing the alpha-2 codes for every country in that continent.
Next using the `dplyr` package, the larger enrolments data frame was transformed
to create a new data frame that contained a count of the number of individuals
enrolled from each continent in each month. Again, using the `lubridate` 
package, another column was added to the data frame that contained the date of
the first day of each of the months for the x-axis positioning. This can be seen
in the 03-A.R file in the munge folder. This data was then plotted using 
`ggplot2` as the stacked bar chart in Figure 3. The code for this plot can be
seen in the eda.R file in the src folder.
``` {r months_continents, fig.align="center", fig.cap="Stacked bar chart showing the number of enrolments from each continent in each month from Ferbruary 2016 to January 2019", echo=FALSE}
months_continents_plot
```
Next, similar to the data preparation in the second cycle, a new data frame was
created by filtering the enrolments from the larger enrolments data frame to
those that fell in the selected two-year time period, except this time the
individuals were also assigned a continent. The learners were then grouped by
the season they enrolled and the continent from which they were detected. A
count was performed for each group. This was then plotted to give the stacked
bar chart in Figure 4.
``` {r seasons_continents, fig.align="center", fig.cap="Stacked bar chart showing the number of enrolments from each continent per season from May 2016 to May 2018", echo=FALSE}
seasons_continents_plot
```
Finally, a data frame was made for each season by filtering the larger 
enrolments data frame to enrolments that fell in the two-year period and that
that were part of the specified season. These were then assigned a continent
which they were then grouped by and a count for each was performed. These four
data frames were then plotted as the pie charts seen in Figure 5 with the 
percentage proportion for each continent in each season being calculated and 
included in the legend for each pie chart.
```{r pies, fig.align="center", fig.cap="Pie charts showing the proportion of enrolments from each continent per season from May 2016 to May 2018", echo=FALSE}
grid.arrange(spring_pie, summer_pie, autumn_pie, winter_pie, ncol=2,
             top=textGrob("Continental Proportion per Season",
                          gp=gpar(fontsize=20,font=1)))
```
## Evaluation
Looking at Figures 3, 4 and 5 it was clear to see that the largest proportion of
learners were situated in Europe (green), followed by Asia (yellow). From this
it was suggested that it could be useful to employ staff that are able to speak
languages commonly spoken in these areas of the Earth for example English,
Mandarin and Hindi to accommodate for the large number of students from
these areas. Focusing on Figure 5, comparing the percentages between the seasons
it could be seen that the proportion of learners from Asia was highest in
summer (25.87%) and lowest in spring (17.55%) so it could be suggested that
it may be sensible to employ more staff who speak languages commonly spoken in 
Asia during summer periods than during spring periods. Additionally, the
proportion of individuals from Africa is largest during the autumn periods 
(16.06% compared with a low of 8.70% in summer periods) so it may be useful
during autumn periods to employ some staff who speak languages commonly spoken 
in Africa such as Arabic. These changes were suggested with the intention of
making the learners more comfortable on the course and potentially enable them 
the opportunity to ask questions in their native languages. Also, if people knew
they were to have this support on the course they may be more inclined to sign 
up to a FutureLearn course and hence increase the number of enrolments. Making
students feel more comfortable and increasing enrolments both fall in line with
FutureLearn's business goals. Moreover, this data could also be used by
FutureLearn to target specific areas at different times when advertising their
courses with knowledge of when people from those areas are more likely to sign 
up which would potentially both increase the number of enrolments, giving them
a larger share of the market and also increase the efficiency of their 
advertising budget, in turn possibly saving them money.

While from these results and the current data, it is possible to determine the times when particular areas take up a larger share of enrolemnts, it is still not feasible to give an exact explanation as to why this may be the case. FutureLearn may want to investigate this more in future analysis however they will still be able to make educated decisions based on finding from this analysis.

Some final limitations of this investigation are that data from only one course
was used so results are limited as the observations from this project may be 
different for other courses, and the length of time over which data was 
collected may not be long enough to accurately predict the seasonal trends. To combat this
in future, data collected from multiple courses over a longer period of time may
be used. Finally, as a recommendation for additional cycles to this project it
might be useful to investigate which individual countries have the largest
proportion of enrolments.

# References
- Shacklock, X., 2016. *From bricks to clicks: The potential of data and analytics in higher education.* London: Higher Education Commission.
- FutureLearn. n.d. FutureLearn: Online Courses and Degrees from Top Universities. [online] Available at: <https://futurelearn.com/> [Accessed 29 November 2021].
- Chapman P, Clinton J, Kerber R, Khabaza T, Reinartz T, Shearer C, Wirth R. CRISP-DM
1.0: Step-by-step data mining guide, 2000.